---
title: "Weight of Fish"
author: "Natasha Lekach"
date: "April 22, 2021"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
library(car)
library(leaps)
```


```{r}
Fishdata=read.csv(file="Fish.csv")
Breamdata<-filter(Fishdata,Species=="Bream")
Smeltdata<-filter(Fishdata,Species=="Smelt")
Perchdata<-filter(Fishdata,Species=="Perch")
```
# Introduction and Methods

Using data from the Fish.csv, we created various models for predicting the weight of three different fish species. 

First, we created new dataframes for each species of fish. Then we removed the species variable, which is a character variable, from the new dataframes to allow leaps to properly create the models.

Next, for each species we drafted various regression models for predicting the weight of each species.  We used the leaps package to draft all the models and then identified the best model with two predictors for predicting weight for each species. After identifying the best model with 2 predictors, we completed an ANOVA to determine if the best 2 predictor model's ability to predict weight is significantly different than the best 1 predictor model.

## Null hypotheses:   
Perch: There is no statistically significant difference between the best 2 predictor model and the best 1 predictor model.They similarly predict the weight.  
Bream: There is no statistically significant difference between the best 2 predictor model and the best 1 predictor model.They similarly predict the weight.  
Smelt: There is no statistically significant difference between the best 2 predictor model and the best 1 predictor model.They  similarly predict the weight.  

## Alternate hypotheses:  
Perch: There is a significant difference between the best 2 predictor model and the best 1 predictor model abilities to predict the weight of perch.  
Bream: There is a significant difference between the best 2 predictor model and the best 1 predictor model abilities to predict the weight of bream.  
Smelt: There is a significant difference between the best 2 predictor model and the best 1 predictor model abilities to predict the weight of smelt.  
 
# PERCH MODEL 

```{r}
filter_Perch<-select(Perchdata,-Species)
```

## Results
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
regsubsets.out <-
    regsubsets(Weight ~ .,
               data = filter_Perch,
               nbest = 1,       # 1 best model for each number of predictors
               nvmax = NULL,    # NULL for no limit on number of variables
               force.in = NULL, force.out = NULL,
               method = "exhaustive")

summary.out <- summary(regsubsets.out)
as.data.frame(summary.out$outmat)
subsets(regsubsets.out,statistic="adjr2",legend="bottomright",main="Figure 1: Adjusted R^2 of Perch Models")
```
  
### Which is best?  

```{r}
which.max(summary.out$adjr2)
```

```{r}
summary.out$which[2,]
```
The best model with 2 predictors is height + width. 
Figure one visually shows that the combination of height and width is the best model with 2 predictors. Additionally, it is the best model overall as it has the highest adjusted r-squared.  The best model with one predictor uses height.
  
```{r}
perch.model <- lm(Weight~Height+Width,data=filter_Perch)
compare.perch<-lm(Weight~Width,data=filter_Perch)
summary(perch.model)

``` 
The best fit 2 predictor model used height + width to predict the weight of perch. This model had an adjusted r squared of 0.94, indicating that 94% of the error encountered by using the mean model is explained when using this model. This model had the highest r-squared overall. The residual standard error was 85.17. This is a pretty high residual standard error.

```{r}
anova(compare.perch,perch.model)
```

We compared the height + width model with the best model that uses one variable, width. Height + width was initially identified as the best model overall and the ANOVA identifies that it is significantly different than model which uses just width - according to the very low p-value.

## Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The adjusted r-squared of this 2 variable perch model is strong and it was identified as the best model. The ANOVA showed a p-value of 0.00112, indicating that the height + width model is significantly different from the best 1 predictor model. There is a very small chance that these results occurred randomly.  

Therefore, we have enough evidence to reject the null hypothesis. If selecting a model to predict weight, fish researchers should use height + weight. 

# BREAM MODEL

```{r}
filter_Bream<-select(Breamdata,-Species)
```

## Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
regsubsets.out1 <-
    regsubsets(Weight ~ .,
               data = filter_Bream,
               nbest = 1,       # 1 best model for each number of predictors
               nvmax = NULL,    # NULL for no limit on number of variables
               force.in = NULL, force.out = NULL,
               method = "exhaustive")

summary.out1 <- summary(regsubsets.out1)
as.data.frame(summary.out1$outmat)
subsets(regsubsets.out1,statistic="adjr2",legend="bottomright",main="Figure 2: Adjusted R^2 of Bream Models")
```

### Which is best?  

```{r}
which.max(summary.out1$adjr2)
```
```{r}
summary.out1$which[2,]
```
The best model overall uses three predictors. The best model with two predictors uses Length2 + height.
From figure 2, we see that the best model with two predictors uses length + height. However, the models with three predictors has a higher adjusted r-squared. The best model with one predictor uses height. 
  
```{r}
BreamModel <- lm(Weight~Length2+Height,data=filter_Bream)
compare.bream<-lm(Weight~Height,data=filter_Bream)
summary(BreamModel)

``` 
The best fit model with two predictors used length2 + height to predict the weight of bream. This model had an adjusted r squared of  0.9356 , indicating that 93.56% of the error encountered by using the mean model is explained when using this model. The residual standard error was 53.1.
  
```{r}
anova(compare.bream,BreamModel)
```

We compared the best model with 2 predictors, length 2 + height, with the best model with 1 predictor. There is a statistically significant difference between these two models. This is determined with the low p-value (0.03638) of this ANOVA.   

## Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

There is a significant difference between the best model with two predictors and the best model with just one predictor. Additionally, figure 2 shows a that the length2 + height model has a higher adjusted r-squared than the height model. There is a very small chance that these results occurred randomly. 

With a significant difference between the models, we have enough evidence to reject the null hypothesis and state that there is a difference between the two models. Bream researchers should select the length2+height model over the height model. 

# SMELT MODEL

```{r}
filter_Smelt<-select(Smeltdata,-Species)
```

## Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
regsubsets.out2 <-
    regsubsets(Weight ~ .,
               data = filter_Smelt,
               nbest = 1,       # 1 best model for each number of predictors
               nvmax = NULL,    # NULL for no limit on number of variables
               force.in = NULL, force.out = NULL,
               method = "exhaustive")

summary.out2 <- summary(regsubsets.out2)
as.data.frame(summary.out2$outmat)
subsets(regsubsets.out2,statistic="adjr2",legend="bottomright",main="Figure 3: Adjusted R^2 of Smelt Models")
```
  
### Which is best?  

```{r}
which.max(summary.out2$adjr2)
```

```{r}
summary.out2$which[2,]
```
Figure three visually shows that the best model with two predictors uses length 2 and width to predict weight. It has a lower adjusted r squared than the models that use three, four and five predictors.   The best model with one predictor uses Length2. The best model overall uses four predictors.
  
```{r}
SmeltModel <- lm(Weight~Length2+Width,data=filter_Smelt)
compare.smelt<-lm(Weight~Length2,data=filter_Smelt)
summary(SmeltModel)

``` 
The best fit model with 2 predicotrs used length2 + width to predict the weight of smelt. This model had an adjusted r squared of  0.9645 , indicating that 96.45% of the error encountered by using the mean model is explained when using this model. The residual standard error was 0.7783. 

```{r}
anova(compare.smelt,SmeltModel)
```

We completed an ANOVA to compare  length2 + width model with the best model which uses just one predictor. The ANOVA returned a p-value of 0.001642, indicating that these two models are significantly different from one another. There is a very small chance that this occurred by chance.

## Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

There is a significant difference between the best model with two predictors and the best model with just one predictor. Figure 3 shows that the length2 + width model has a higher adjusted r-squared than the length2 model. There is a very small chance that these results occurred randomly. 

With a significant difference between the models, we have enough evidence to reject the null hypothesis and state that there is a difference between the models. Smelt researchers should select the length2+width model over that just length2 model. 

Therefore, we have enough evidence to reject the null hypothesis. One should note that this dataset was only made up of 14 observations of smelt. Maybe if the dataset was larger, we would have been able to create a better model.