---
title: "Weight of Fish"
author: "Natasha Lekach"
date: "April 22, 2021"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
```


```{r}
Fishdata=read.csv(file="Fish.csv")
Breamdata<-filter(Fishdata,Species=="Bream")
Smeltdata<-filter(Fishdata,Species=="Smelt")
Perchdata<-filter(Fishdata,Species=="Perch")
```

# PERCH MODEL 
## Introduction to PERCH Model

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Using data from ________________, we will draft various regression models for predicting the weight of perch. After drafting the models, we will determine which is the best model for predicting the weight of this species of fish. 

Null hypotheses: 
The various perch models are not statistically significantly different from each other. They all similarly predict weight.

Alternate hypotheses:
There is a statistically significant difference in at least one of the perch models' ability to predict weight.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We used the leaps package to create all the models and then identify the best models for predicting perch weight. After identifying the best model, we will complete an ANOVA to determine if the best model is significantly better at predicting weight.  

To create these models, we first cleaned up the data by isolating just the perch data from the larger Fishdata dataframe. Then we removed the character variable of species to allow leaps to properly create the models.  

```{r}
filter_Perch<-select(Perchdata,-Species)
```

## Results

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
library(leaps)
regsubsets.out <-
    regsubsets(Weight ~ .,
               data = filter_Perch,
               nbest = 1,       # 1 best model for each number of predictors
               nvmax = NULL,    # NULL for no limit on number of variables
               force.in = NULL, force.out = NULL,
               method = "exhaustive")

summary.out <- summary(regsubsets.out)
as.data.frame(summary.out$outmat)
library(car)
subsets(regsubsets.out,statistic="adjr2",legend="bottomright",main="Figure 1: Adjusted R^2 of Perch Models")
```

##Which is best?  

```{r}
which.max(summary.out$adjr2)
```
Figure one visually shows that the combination and height and width have the highest adjusted R squared. Above, we confirm that option number 2, height and width, are the best predictor of weight.   

```{r}
summary.out$which[3,]
```

```{r}
best.model <- lm(Weight~Height+Width,data=filter_Perch)
fullup<-lm(Weight~.,data=filter_Perch)
summary(best.model)

``` 
```{r}
anova(fullup,best.model)
```

We compared the best model, height plus width, with a model which uses all the variables. While height plus width was initially identified as the best model, there is no statistically significant difference from the other models. This is determined with the high p-value of this ANOVA.  

## Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

# BREAM MODEL
## Introduction to BREAM Model

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Using data from ________________, we will draft various regression models for predicting the weight of bream. After drafting the models, we will determine which is the best model for predicting the weight of this species of fish. 
Each model will use two predictor variables. 
To create these models, we cleaned up the data by isolating just the bream data from the larger Fishdata dataframe. 

Null hypotheses: 
The various bream models are not statistically significantly different from each other. They all similarly predict weight.

Alternate hypotheses:
There is a statistically significant difference in at least one of the bream models' ability to predict weight.


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


## Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

# SMELT MODEL
## Introduction to SMELT model

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Using data from ________________, we will draft various regression models for predicting the weight of smelt. After drafting the models, we will determine which is the best model for predicting the weight of this species of fish. 
Each model will use two predictor variables. 
To create these models, we cleaned up the data by isolating just the smelt data from the larger Fishdata dataframe. 

Null hypotheses: 
The various smelt models are not statistically significantly different from each other. They all similarly predict weight.

Alternate hypotheses:
There is a statistically significant difference in at least one of the smelt models' ability to predict weight.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


## Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
